The Error Estimate of Force Calculation in the Inhomogeneous Molecular Systems

### 2011-12-15 shortRange.inhomo/writeDown/main.tex
revise according to Christof

\documentclass[aps,pre,preprint]{revtex4-1}
% \documentclass[aps,pre,twocolumn]{revtex4-1}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

\begin{abstract}
\end{abstract}

@@

In this paper, for simplicity,
we will consider one-component systems, but
% we only consider the single-component system. Because
% this assumption simplifies the discussion, and
the error estimates for
the one-component system can be straightforwardly extended to the
multi-component systems.  we consider
force error in the present paper, because the
applications are based on molecular dynamics simulation.
% Since the error estimate developed for the one-component system can be
% directly extended to multi-component system without substantial
% difficulty, the discussion in this paper will focus on one-component
% system for symplicity. 
Suppose the system is composed by $N$ identical particles located at
$\v r_1, \v r_2, \cdots, \v r_N$ with periodic boundary condition.
These particles are interacting via a short-range pairwise interaction
$u(\v r)$ with a cut-off radius $r_c > 0$. The corresponding force is
denoted by $\v f(\v r)$. 
% An infinite cut-off radius will lead to
% a precise calculation of interaction, but it is impractical.
% The error of computing interaction is mainly introduced by a finit
% value of the cut-off radius.
The force error mainly stems from the neglected interaction outside
the cut-off radius, so it is convenient to consider the complementary
of the cut-offed force that is defined by
\begin{align}
  \v f^c(\v r) =
  \left\{
  \begin{array}{ll}
    0, & \vert\v r\vert\leq r_c; \\
    \v f(\v r), & \vert\v r\vert > r_c.
  \end{array}
  \right.
\end{align}
We measure the force error at an arbitrary position in the system,
saying $\v r$, by adding a testing particle to that position.  We
assume all particles in the system exert forces on the testing
particle, whereas the testing particle does NOT exert any force on the
system.  The exact force and the computed force (or cut-offed force)
exerted by the system on the testing particle are denoted by ${\v
  F}(\v r)$ and $\tilde{\v F}(\v r)$, respectively.  The difference
between them, namely the \emph{error force}, is denoted by $\Delta\v
F(\v r) = \v F(\v r) - \tilde{\v F}(\v r)$, which can be expressed by
\begin{align}
  \Delta\v F(\v r) = \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n).
\end{align}
Where $\v n$ is the box vector, so the summation over $\v n$ indicates
that all periodic images of the system are considered. In a
homogeneous system, the surrounding of any particle is isotropic, so
the \emph{mean error force} vanishes: $\langle\Delta\v F(\v r)\rangle
= 0$. In contrast, the mean error force is not necessarily zero in an
inhomogeneous system:
\begin{align} \label{eqn:tmp3} %\nonumber
  \langle\Delta\v F(\v r)\rangle
  &=
  \big\langle
  \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n)
  \big\rangle 
  =
  \int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r',
\end{align}
where $\rho(\v r)$ denotes the particle number density at the position
$\v r$, and is periodically extended to $\mathbb R^3$. Notice that the
error force implicitly depends on the cut-off radius used for the
interaction calculation.

@@

The naive calculation of the convolutions in the error estimate \eqref{eqn:tmp5} requires
a computational cost of $\mathcal O(N^2)$. To estimate the force error
on the fly, the computational cost of the estimate should be reduced to
at least $\mathcal O(N\log N)$.  This is achieved by  using the fast Fourier transform. We
assume the simulation box is uniformly divided in to small bins.  The
positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{K_1}\v a_1 + 
  \frac{i_2}{K_2}\v a_2 + 
  \frac{i_3}{K_3}\v a_3,\qquad 0\leq i_\alpha< K_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $K_\alpha$ is the number of
division on each direction. In the reciprocal space, a corresponding
lattice is set up:
\begin{align}
  \v k_{k_1, k_2, k_3} & =
  k_1\v a_1^\ast +
  k_2\v a_2^\ast +
  k_3\v a_3^\ast, \qquad 0\leq k_\alpha<K_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
then, a Fourier expansion can be obtained:
\begin{align}\nonumber
  \mathcal E^2_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & =
  \frac1V\sum_{k_1=-\infty}^\infty\sum_{k_2=-\infty}^\infty\sum_{k_3=-\infty}^\infty
  [{\mathcal E}^2_{\textrm{homo}}]^{\wedge}(\v k_{k_1,k_2,k_3})
  % e^{2\pi i\v k_{k_1,k_2,k_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & =
  \frac1V\sum_{k_1=0}^{K_1}\sum_{k_2=0}^{K_2}\sum_{k_3=0}^{K_3}
  \hat{K}_{\textrm{homo}}(\v k'_{k_1,k_2,k_3})\:
  \hat{\rho}(\v k_{k_1,k_2,k_3})
  % e^{2\pi i\v k_{k_1,k_2,k_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg].
\end{align}
Here $K_{\textrm{homo}}(\v r)$ denotes $\vert\v f^c(\v r)\vert^2$.
The wedge notation $\wedge$ means Fourier transform. The prime on $\v k_{k_1,k_2,k_3}$ means
that the periodic image $k_\alpha - K_\alpha$ should use instead of
$k_\alpha$ when $k_\alpha \geq K_\alpha/2$.
The second equation in \eqref{eqn:tmp13}
holds due to the fact:
$[\mathcal E^2]^\wedge = [K_{\textrm{homo}}\ast\rho]^\wedge = \hat{K}_{\textrm{homo}} \times \hat\rho$. The Fourier transform of $\rho$ and $K_{\textrm{homo}}$  are
\begin{align}
  \hat\rho(\v k_{k_1,k_2,k_3})
  & =
  \frac{V}{K_1K_2K_3}
  \sum_{i_1=0}^{K_1}\sum_{i_2=0}^{K_2}\sum_{i_2=0}^{K_2}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg],
\end{align}
and
\begin{align}\label{eqn:f-kernel-homo}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v s)e^{-2\pi i\v k\cdot\v s}\,\d d\v s.
\end{align}
Most of the short-range
interactions are isotropic, i.e. $K_{\textrm{homo}}(\v r) =
K_{\textrm{homo}}(r)$, then the three-dimensional integral in
Eqn. \eqref{eqn:f-kernel-homo} can be simplified to a one-dimensional
integral:
\begin{align}\label{eqn:kernel-inte}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  % \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  % =
  \int_{r_c}^\infty \frac{2 s}k\, K_{\textrm{homo}}(s) \sin(2\pi ks)\,\d ds,
\end{align}
which can be calculated by numerical
integration from $r_c$ to a large upper bound, saying $R_c$. Perfect
convergence can be achieved by using $R_c = 30\sigma$ throughout the
paper. Since integral \eqref{eqn:kernel-inte} is independent of the
density of the system, it is calculated only once and stored.

@@

A basic observation in an inhomogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be more than one order of
magnitude
% in one of the cases we studied,
(see Fig. \ref{fig:tmp2}). Traditionally, a very large cut-off radius is
used to reduce the maximum error in the system to reach convergent
simulation results~\cite{ismail2007application}. Therefore, a huge
amount of computational effort is wasted in the low-error region. The
idea of the adaptive cut-off radius simulation is to use larger
cut-off radius for particles in the high-error region to control the
force error, while to use smaller cut-off radius for particles in the
low-error region to save computational cost.  The effectiveness of this
method relies on a deliberated cut-off choice for each particle so
that the force error is uniformly distributed across the simulation
region.  Unfortunately, the computational expense of a naive
implementation of this method is at least $\mathcal O(N^2\log N)$.
Because in order to determine the cut-off radius, we should estimate the RMS
force error at least once for each particle, and the computational
cost of each force error estimate is $\mathcal O(N \log N)$.

@@

To make the adaptive cut-off method feasible, we assume:
\begin{enumerate}
\item The space is uniformly divided in to small bins, the size of which is the
  same as those mentioned in the section \ref{sec:tmp1.3}.  The
  particles falling in the same bin are assigned the same cut-off
  radius. In other words, the cut-off radius is chosen for each bin
  rather than for each particles.
  % The cut-off
  % radius is assumed to be constant within a bin, no matter the
  % position of the partile.
\item The cut-off radius can only be chosen from a discrete series of
  monotonically increasing values between a maximum 
  $r_c^{\textrm{max}}$ and a minimum 
  $r_c^{\textrm{min}}$: $\{\,r_c^{\textrm{min}} = r_c^0 < r_c^1 <
  \cdots < r_c^{M-1} <r_c^M = r_c^{\textrm{max}}\}$.  We call it the
  \emph{candidate set} of the cut-off radius.  For convenience but not
  necessarily, the cut-off radii in the candidate set is uniformly increasing, i.e.
  \begin{align}
    r_c^m = r_c^0 + r_c^{\textrm{step}} m, \quad 0 \leq m \leq M.
  \end{align}
  For example, a possible candidate set is $\{2.5\,\sigma,\
  3.0\,\sigma,\ 3.5\,\sigma,\ 4.0\,\sigma,\ \cdots,\ 10.0\,\sigma\}$,
  with $r_c^{\textrm{max}} = 10.0\sigma$, $r_c^{\textrm{min}} =
  2.5\sigma$ and $r_c^{\textrm{step}} = 0.5 \sigma$.
\end{enumerate}
For each cut-off radius in the candidate set, the force error
$\mathcal E(\v r)$ is calculated by the estimate Eqn.~\eqref{eqn:tmp5}.
Notice the error is calculated by the
fast algorithm described in section~\ref{sec:tmp1.3}, so it is
defined on the bins.  For a predetermined error control level
$\mathcal E_{\textrm{C}}$, a satisfactory cut-off radius for a bin is
chosen such that it is the smallest one satisfying the precision
constraint $\mathcal E(\v r_{i_1,i_2,i_3}) \leq \mathcal
E_{\textrm{C}}$. The computational cost of this algorithm is $\mathcal
O(MN\log N)$.  The idea of this algorithm is based on the observation
that the cut-off radius may not be calculated at a very high
precision, a slightly (at most $r_c^{\textrm{step}}$) larger
cut-off radius will not waste too much computational cost.  Due to the
thermodynamic fluctuation of the particle density, the resulting cut-off
distribution $r_c(\v r_{i_1,i_2,i_3})$ (as a function of bin
position) needs to be refined by the following scheme:
\begin{align}\label{eqn:rc-refine}
  r^{(1)}_c(\v r_{i_1,i_2,i_3}) = \max_{j_\alpha\in I}\,r_c(\v r_{i_1+j_1,i_2+j_2,i_3+j_3}),
  \quad I = \{-1, 0, 1\},
\end{align}
so that the fluctuating large-error regions are safely covered by bins
with large cut-off radius.

@@

The idea of the long-range force correction is very similar to those
proposed by earlier works~\cite{guo1997long, mecke1997molecular,
  janecek2006long, shen2007comparative}. In those works, the density
profile is assumed to be inhomogeneous along one direction only, and the
system is subdivided into slabs perpendicular to that direction to
calculate the energy/force correction. The advantage of our force
correction is that it has no geometric restriction on the density
profile of the system, so can be used in a wider range of
inhomogeneous simulations.

@@

The direct simulation of the liquid-vapor equilibrium is a very
typical inhomogeneous system, which is widely used to study the phase
coexistence properties, for example, the equilibrium liquid/vapor
density, and the surface tension. The system studied in the present
paper contains 16,000 identical standard Lennard-Jones particles that are
separated into liquid phase in the center of the box and vapor phase
surrounded (see Fig.~\ref{fig:tmp1}).  The conventional Lennard-Jones
unit system is employed: the unit of length, energy, mass and time are
denoted by $\epsilon$, $\sigma$, $m$ and $\tau$ ($\tau =
\sigma\sqrt{m/\varepsilon}$), respectively.  All quantities are
rescaled by the Lennard-Jones units and added the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$,
$P^\ast = P\cdot \sigma^3 / \epsilon$ and $\mathcal E^\ast = \mathcal
E\cdot\sigma/\varepsilon$.  The MD time step {is} $\Delta t^\ast =
0.005$. The simulations last for $3\times 10^6$ time steps. The first
$1\times 10^6$ steps are discarded. The quantities of interest are
sampled every 100 time steps. The blocking average
method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data with 95\%
confidence level. The NVT ensemble is generated by the Nos\`e-Hoover
thermostat~\cite{nose1984molecular, hoover1985canonical}.  The size of
the simulation box is $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 21\times 21$, which permits a maximum cut-off radius of
$r_c^\ast = 10$, and eliminates the finite size effect reported in
literatures \cite{chen1995area, orea2005oscillatory,
  biscay2009calculation}.
The system is divided into bins of size $1\times1\times1$ to perform the fast error
estimate method described in section \label{sec:tmp1.3}.
In all simulations, the maximum cut-off
radius considered is $10$, but the reference results for comparison
are obtained by simulations carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$.

@@

% The Fig. \ref{fig:tmp2} inspires us to develop the adaptive cut-off
% method: in the bulk regions, the error is much smaller, therefore, it
% is a waste to use the same cut-off radius as the interfacial
% regions.
As mentioned before, the key idea of the adaptive cut-off simulation
is to use smaller cut-off
radius in the bulk regions, so that the RMS force error is uniformly
distributed over the simulation region. In Fig. \ref{fig:tmp3}, we
present the adapted cut-off radius by a solid red line and all errors
by the same notations as in the Fig. \ref{fig:tmp2}. The control error
$\mathcal E^\ast_{\textrm{C}} = 0.0045$ is the same as the maximum
error in the uniform cut-off simulation with $r_c^\ast=7.5$. The
cut-off distribution is calculated every 20,000 time steps, and is
refined twice by Eqn.~\eqref{eqn:rc-refine}. In the interfacial
regions, the same cut-off radius as the uniform cut-off simulation,
namely $7.5$, is used.  In the
liquid and vapor regions, the cut-off radii are $4.75$ and $3.5$,
respective, which is notably reduced. Since the computational cost
grows with $\mathcal O(r_c^3)$, the adaptive cut-off saves 75 \% and
90 \% computational cost in the liquid and vapor bulk regions,
respectively.  The error is uniformly distributed, except in the
liquid region where the real error is somewhat lower.
In contrast with the uniform
cut-off case, i.e. Fig. \ref{fig:tmp2}, the error estimate for the
vapor region is very sharp. The similar phenomenon is observed in
Ref.~\cite{wang2011}. The same as the uniform cut-off simulation, the
homogeneity error dominates in the bulk regions, while the
inhomogeneity error dominates in the interfacial region.

>>>>

\documentclass[aps,pre,preprint]{revtex4}
% \documentclass[aps,pre,twocolumn]{revtex4-1}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

\begin{abstract}
  \recheck{
    In the present paper, we develop an accurate error estimate of the
    non-bonded short-range interaction for the inhomogeneous molecular
    systems.
    The root mean square force error is proved to be decomposed into
    three additive parts: the homogeneity error, the inhomogeneity error
    and the correlation error.
    % The homogeneity error is dominant
    % in the bulk regions.
    The magnitude of the inhomogeneity error, which
    is dominant in the interfacial regions, 
    can be more than one order of magnitude larger than the homogeneity
    error.
    % This is the reason why a direct simulation may waste
    % a huge amount of computational effort.
    This is the reason why a standard simulation with fixed
    cut-off radius is either less accurate if the cut-off is too small,
    or wastes considerable computational effort if the cut-off is too large.
    Therefore, based on the error estimate, the adaptive cut-off
    and long-range force correction methods are proposed to
    boost the efficiency and accuracy of the simulation, respectively.
    The
    effectiveness of the proposed methods are demonstrated by
    molecular dynamics 
    simulations of the liquid-vapor equilibrium and the nanoscale particle
    collision. Different roles of the homogeneity error and inhomogeneity
    error are also discussed.
  }
\end{abstract}

@@

In this paper, for simplicity,
we will consider one-component systems, but
% we only consider the single-component system. Because
% this assumption simplifies the discussion, and
the error estimates for
the one-component system can be straightforwardly extended to the
multi-component systems.  We consider
the force error in the present paper, because the
applications are based on molecular dynamics simulations.
% Since the error estimate developed for the one-component system can be
% directly extended to multi-component system without substantial
% difficulty, the discussion in this paper will focus on one-component
% system for symplicity. 
Suppose the system is composed by $N$ identical particles located at
$\v r_1, \v r_2, \cdots, \v r_N$ with periodic boundary condition.
These particles are interacting via a short-range pairwise interaction
$u(\v r)$ with a cut-off radius $r_c > 0$. The corresponding force is
denoted by $\v f(\v r)$. 
% An infinite cut-off radius will lead to
% a precise calculation of interaction, but it is impractical.
% The error of computing interaction is mainly introduced by a finit
% value of the cut-off radius.
The force error mainly stems from the neglected interaction outside
the cut-off radius, so it is convenient to consider the complementary
of the cut-offed force that is defined by
\begin{align}
  \v f^c(\v r) =
  \left\{
  \begin{array}{ll}
    0, & \vert\v r\vert\leq r_c; \\
    \v f(\v r), & \vert\v r\vert > r_c.
  \end{array}
  \right.
\end{align}
We measure the force error at an arbitrary position in the system,
saying $\v r$, by adding a testing particle to that position.  We
assume all particles in the system exert forces on the testing
particle, whereas the testing particle does NOT exert any force on the
system.  The exact force and the computed force (or cut-offed force)
exerted by the system on the testing particle are denoted by ${\v
  F}(\v r)$ and $\tilde{\v F}(\v r)$, respectively.  The difference
between them, namely the \emph{error force}, is denoted by $\Delta\v
F(\v r) = \v F(\v r) - \tilde{\v F}(\v r)$, which can be expressed by
\begin{align}
  \Delta\v F(\v r) = \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n).
\end{align}
Where $\v n$ is the box vector, so the summation over $\v n$ indicates
that all periodic images of the system are considered. In a
homogeneous system, the surrounding of any particle is isotropic, so
the \emph{mean error force} vanishes: $\langle\Delta\v F(\v r)\rangle
= 0$. In contrast, the mean error force is not necessarily zero in an
inhomogeneous system:
\begin{align} \label{eqn:tmp3} %\nonumber
  \langle\Delta\v F(\v r)\rangle
  &=
  \big\langle
  \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n)
  \big\rangle 
  =
  \int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r',
\end{align}
where $\rho(\v r)$ denotes the particle number density at the position
$\v r$, and is periodically extended to $\mathbb R^3$. Notice that the
error force implicitly depends on the cut-off radius used for the
interaction calculation.

@@

The naive calculation of the convolutions in the error estimate \eqref{eqn:tmp5} requires
a computational cost of $\mathcal O(N^2)$. To estimate the force error
on the fly, the computational cost of the estimate should be reduced to
at least $\mathcal O(N\log N)$.  This is achieved by  using the fast Fourier transform. We
assume the simulation box is uniformly divided in to small bins.  The
positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{M_1}\v a_1 + 
  \frac{i_2}{M_2}\v a_2 + 
  \frac{i_3}{M_3}\v a_3,\qquad 0\leq i_\alpha< M_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $M_\alpha$ is the number of
divisions on each direction. In the reciprocal space, a corresponding
lattice is set up:
\begin{align}
  \v k_{m_1, m_2, m_3} & =
  m_1\v a_1^\ast +
  m_2\v a_2^\ast +
  m_3\v a_3^\ast, \qquad 0\leq m_\alpha<M_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
then, a Fourier expansion can be obtained:
\begin{align}\nonumber
  \mathcal E^2_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & =
  \frac1V\sum_{m_1=-\infty}^\infty\sum_{m_2=-\infty}^\infty\sum_{m_3=-\infty}^\infty
  [{\mathcal E}^2_{\textrm{homo}}]^{\wedge}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & =
  \frac1V\sum_{m_1=0}^{M_1}\sum_{m_2=0}^{M_2}\sum_{m_3=0}^{M_3}
  \hat{K}_{\textrm{homo}}(\v k'_{m_1,m_2,m_3})\:
  \hat{\rho}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg].
\end{align}
Here $K_{\textrm{homo}}(\v r)$ denotes $\vert\v f^c(\v r)\vert^2$.
The wedge notation $\wedge$ means the Fourier transform. The prime on $\v k_{m_1,m_2,m_3}$ means
that the periodic image $m_\alpha - M_\alpha$ should use instead of
$m_\alpha$ when $m_\alpha \geq M_\alpha/2$.
The second equation in \eqref{eqn:tmp13}
holds due to the fact:
$[\mathcal E^2]^\wedge = [K_{\textrm{homo}}\ast\rho]^\wedge = \hat{K}_{\textrm{homo}} \times \hat\rho$. The Fourier transform of $\rho$ and $K_{\textrm{homo}}$  are
\begin{align}
  \hat\rho(\v k_{m_1,m_2,m_3})
  & =
  \frac{V}{M_1M_2M_3}
  \sum_{i_1=0}^{M_1}\sum_{i_2=0}^{M_2}\sum_{i_3=0}^{M_3}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg],
\end{align}
and
\begin{align}\label{eqn:f-kernel-homo}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v s)e^{-2\pi i\v k\cdot\v s}\,\d d\v s.
\end{align}
Most of the short-range
interactions are isotropic, i.e. $K_{\textrm{homo}}(\v r) =
K_{\textrm{homo}}(r)$, then the three-dimensional integral in
Eqn. \eqref{eqn:f-kernel-homo} can be simplified to a one-dimensional
integral:
\begin{align}\label{eqn:kernel-inte}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  % \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  % =
  \int_{r_c}^\infty \frac{2 s}k\, K_{\textrm{homo}}(s) \sin(2\pi ks)\,\d ds,
\end{align}
which can be calculated by numerical
integration from $r_c$ to a large upper bound, saying $r_u$. \recheck{Perfect
convergence is achieved by using $r_u = 30\sigma$
for the dispersion energy $-4\epsilon (\sigma / r)^6$, where $\epsilon$
and $\sigma$ are the energy and length units.}
Since integral~\eqref{eqn:kernel-inte} is independent of the
density of the system, it is calculated only once and stored.

@@

A basic observation in an inhomogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be more than one order of
magnitude
% in one of the cases we studied,
(see Fig. \ref{fig:tmp2}). Traditionally, a very large uniform cut-off radius is
used to reduce the maximum error in the system to reach convergent
simulation results~\cite{ismail2007application}.
\recheck{
  However, since the computational effort scales as the
  cube of the cut-off radius, a slightly larger  cut-off
  will notably increase the computational effort.
  For example, increasing the cut-off radius
  from $4\sigma$ to $5\sigma$ will cost
  twice the computational effort.}
Therefore, a huge
amount of computational effort is wasted in the low-error region. The
idea of the adaptive cut-off radius simulation is to use larger
cut-off radius for particles in the high-error region to control the
force error, while to use smaller cut-off radius for particles in the
low-error region to save computational cost.  The effectiveness of this
method relies on a deliberated cut-off choice for each particle so
that the force error is uniformly distributed across the simulation
region.
 In order to determine the cut-off radius, we should estimate the RMS
force error at least once for each particle, and the computational
cost of each force error estimate is $\mathcal O(N \log N)$,
so the total computational expense of 
this method is at least $\mathcal O(N^2\log N)$, which is
prohibitive to the on-the-fly application.

@@

To make the adaptive cut-off method feasible, we assume:
\begin{enumerate}
\item The space is uniformly divided in to small bins, the size of which is the
  same as those mentioned in the section \ref{sec:tmp1.3}.  The
  particles falling in the same bin are assigned the same cut-off
  radius. In other words, the cut-off radius is chosen for each bin
  rather than for each particles.
  % The cut-off
  % radius is assumed to be constant within a bin, no matter the
  % position of the partile.
\item The cut-off radius can only be chosen from a discrete series of
  monotonically increasing values between a maximum 
  $r_c^{\textrm{max}}$ and a minimum 
  $r_c^{\textrm{min}}$: $\{\,r_c^{\textrm{min}} = r_c^0 < r_c^1 <
  \cdots < r_c^{M-1} <r_c^M = r_c^{\textrm{max}}\}$.  We call it the
  \emph{candidate set} of the cut-off radius.  For convenience but not
  necessarily, the cut-off radii in the candidate set is uniformly increasing, i.e.
  \begin{align}
    r_c^m = r_c^0 + r_c^{\textrm{step}} m, \quad 0 \leq m \leq M.
  \end{align}
  For example, a possible candidate set is $\{2.5\,\sigma,\
  3.0\,\sigma,\ 3.5\,\sigma,\ 4.0\,\sigma,\ \cdots,\ 10.0\,\sigma\}$,
  with $r_c^{\textrm{max}} = 10.0\sigma$, $r_c^{\textrm{min}} =
  2.5\sigma$ and $r_c^{\textrm{step}} = 0.5 \sigma$.
\end{enumerate}
For each cut-off radius in the candidate set, the force error
$\mathcal E(\v r)$ is calculated by the estimate Eqn.~\eqref{eqn:tmp5}.
Notice the error is calculated by the
fast algorithm described in section~\ref{sec:tmp1.3}, so it is
defined on the bins.  For a predetermined error control level
$\mathcal E_{\textrm{C}}$, a satisfactory cut-off radius for a bin is
chosen such that it is the smallest one satisfying the precision
constraint $\mathcal E(\v r_{i_1,i_2,i_3}) \leq \mathcal
E_{\textrm{C}}$. The computational cost of this algorithm is $\mathcal
O(MN\log N)$.  The idea of this algorithm is based on the observation
that the cut-off radius may not be calculated at a very high
precision.
% a slightly (at most $r_c^{\textrm{step}}$) larger
% cut-off radius
\recheck{
  Because if $r_c^{\textrm{step}}$ is sufficiently small,
  increasing the cut-off radius by no more than $r_c^{\textrm{step}}$
will not waste too much computational cost.
Here we recommend that $r_c^{\textrm{step}}$ should be no more than
0$.25\sigma$ that increases
the computational effort by 27\% at $r_c = 3.0\sigma$.}
Due to the
thermodynamic fluctuation of the particle density, the resulting cut-off
distribution $r_c(\v r_{i_1,i_2,i_3})$ (as a function of bin
position) needs to be refined by the following scheme:
\begin{align}\label{eqn:rc-refine}
  r^{(1)}_c(\v r_{i_1,i_2,i_3}) = \max_{j_\alpha\in I}\,r_c(\v r_{i_1+j_1,i_2+j_2,i_3+j_3}),
  \quad I = \{-1, 0, 1\},
\end{align}
so that the fluctuating large-error regions are safely covered by bins
with large cut-off radius.

@@

The idea of the long-range force correction is very similar to those
proposed by earlier works~\cite{guo1997long, mecke1997molecular,
  janecek2006long, shen2007comparative}. In those works, the density
profile is assumed to be inhomogeneous along one direction, and uniform
on the other two directions.
The
system is subdivided into slabs perpendicular to the inhomogeneous direction to
calculate the energy/force correction. The advantage of our force
correction is that it has \emph{no} uniformity restriction on the density
profile of the system, so can be used in a wider range of
inhomogeneous simulations.

@@

The direct simulation of the liquid-vapor equilibrium is a very
typical inhomogeneous system, which is widely used to study the phase
coexistence properties, for example, the equilibrium liquid/vapor
density, and the surface tension. The system studied in the present
paper contains 16,000 identical standard Lennard-Jones particles that are
separated into liquid phase in the center of the periodic box and vapor phase
surrounded (see Fig.~\ref{fig:tmp1}).  The conventional Lennard-Jones
unit system is employed: the unit of length, energy, mass and time are
denoted by $\epsilon$, $\sigma$, $m$ and $\tau$ ($\tau =
\sigma\sqrt{m/\varepsilon}$), respectively.  In the rest of this paper,
all quantities are
rescaled by the Lennard-Jones units and added the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$,
$P^\ast = P\cdot \sigma^3 / \epsilon$ and $\mathcal E^\ast = \mathcal
E\cdot\sigma/\varepsilon$.  The MD time step {is} $\Delta t^\ast =
0.005$. The simulations last for $3\times 10^6$ time steps. The first
$1\times 10^6$ steps are discarded. The quantities of interest are
sampled every 100 time steps. The blocking average
method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data with 95\%
confidence level. The NVT ensemble is generated by the Nos\`e-Hoover
thermostat~\cite{nose1984molecular, hoover1985canonical}.  The size of
the simulation box is $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 21\times 21$, which permits a maximum cut-off radius of
$r_c^\ast = 10$, and eliminates the finite size effect reported in
literatures \cite{chen1995area, orea2005oscillatory,
  biscay2009calculation}.
The system is divided into bins of size $1\times1\times1$ to perform the fast error
estimate method described in section \ref{sec:tmp1.3}.
In all simulations, the maximum cut-off
radius considered is $10$, but the reference results for comparison
are obtained by simulations carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$. \recheck {We find that (not shown here)
  the bigger system leads to the same equilibrium
  quantities as the smaller system at the same cut-off radius, so
  it is reasonable to
  use it to produce the better converged data at $r_c^\ast=13$ and compare them with
  those obtained in the smaller system.
}

@@

% The Fig. \ref{fig:tmp2} inspires us to develop the adaptive cut-off
% method: in the bulk regions, the error is much smaller, therefore, it
% is a waste to use the same cut-off radius as the interfacial
% regions.
As mentioned before, the key idea of the adaptive cut-off simulation
is to use smaller cut-off
radius in the bulk regions, so that the RMS force error is uniformly
distributed over the simulation region. In Fig. \ref{fig:tmp3}, we
present the adapted cut-off radius by a solid red line and all errors
by the same notations as in the Fig. \ref{fig:tmp2}. The control error
$\mathcal E^\ast_{\textrm{C}} = 0.0045$ is almost the same as the maximum
error in the uniform cut-off simulation with $r_c^\ast=7.5$.
\recheck{
  In the rest of the paper, when we mention that ``the cut-off of
  the adaptive cut-off method is $r_c^\ast$'', we mean the
  control error is the maximum error of a uniform cut-off simulation
  using that cut-off $r_c^\ast$. In all the cases considered by
  the present paper, the maximum of
  the adapted cut-off radius is almost the same as $r_c^\ast$,
  and is sometimes marginally larger (no more than $r_c^\textrm{step}$).
}
The
cut-off distribution is calculated every 20,000 time steps, and is
refined twice by Eqn.~\eqref{eqn:rc-refine}. In the interfacial
regions, the same cut-off radius as the uniform cut-off simulation,
namely $7.5$, is used.  In the bulk
liquid and vapor regions, the cut-off radii are $4.75$ and $3.5$,
which reduce the computational cost
by a factor of 4 and 10, respectively.
The error is uniformly distributed, except in the
liquid region where the real error is somewhat lower.
In contrast with the uniform
cut-off case, i.e. Fig. \ref{fig:tmp2}, the error estimate for the
vapor region is very sharp.
% The similar phenomenon is observed in
% Ref.~\cite{wang2011}.
The same as the uniform cut-off simulation, the
homogeneity error dominates in the bulk regions, while the
inhomogeneity error dominates in the interfacial region.
